---
title: "Reproducibility Report: Racial/Ethnic Disparities in Driver Gene Mutations"
subtitle: "MSK-CHORD 2024 Analysis | BIOSTAT 625 / PUBHLTH 614"
author: "Kai Wei"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: kate
    code_folding: show
    number_sections: true
    df_print: kable
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  message   = FALSE,
  warning   = FALSE,
  fig.align = "center",
  fig.width = 8,
  fig.height = 5
)
```

---

# Overview

This document reports how the MSK-CHORD 2024 driver-gene mutation analysis
was designed and executed according to reproducible research principles. The
five tools from Dempsey & Morrison (2023) are each addressed in a dedicated
section: (1) file organisation and naming, (2) version control, (3) data
manipulation practices, (4) programming habits, and (5) pipeline design.

**Research question.** Do driver gene mutation frequencies differ
across self-reported racial/ethnic groups in non-small cell lung cancer
(NSCLC), and how do these differences associate with overall survival, after
adjusting for smoking history, disease stage, histologic subtype, sex, and
age?

**Dataset.** MSK-CHORD 2024, downloaded from cBioPortal
(https://www.cbioportal.org/study/summary?id=msk_chord_2024).
The cohort includes 25,040 patients across multiple cancer types; 7,775
NSCLC patients with complete covariate data are used for the primary analysis.

---

# Tool 1 — File Organisation and Naming

## Rationale

Dempsey & Morrison recommend a deliberate directory structure so that any
collaborator (including your future self) can navigate the project without
asking for help. File names should be descriptive and date-stamped where
appropriate. Numeric prefixes on script names encode execution order,
removing ambiguity about which script to run first.

## Directory structure

The project root is `msk_chord_analysis/`. Every path in every script
is constructed with `here::here()` so the project is fully portable —
moving the root folder or running on a different operating system requires
no path edits.

```
msk_chord_analysis/
│
├── R/
│   ├── 00_setup.R               ← packages, seeds, global constants
│   ├── 01_data_processing.R     ← download, clean, impute
│   ├── 02_descriptive.R         ← Table 1, prevalence bar plots
│   ├── 03_chi_square_fdr.R      ← Fisher/chi-square + BH-FDR
│   ├── 04_logistic_regression.R ← multivariable logistic, forest plots
│   ├── 05_survival_analysis.R   ← KM curves, Cox PH + interaction
│   ├── 06_replication_cohorts.R ← CRC and Breast cohort replication
│   └── 07_run_all.R             ← master pipeline (sources 01–06 in order)
│
├── data/
│   └── raw/                     ← raw TSV / mutation TXT from cBioPortal
│                                   (.gitignore excludes this directory)
│
├── output/
│   ├── figures/                 ← PDFs produced by ggplot2 / survminer
│   └── tables/                  ← CSVs produced by each analysis step
│
├── reproducibility_report.Rmd   ← this document
├── README.md
└── .gitignore
```

The numeric prefix convention (00, 01, …, 07) means every file manager
and `ls` output displays scripts in execution order without any additional
documentation.

## Verifying the structure from R

The chunk below inspects the actual project tree at knit time.

```{r file-structure}
library(here)
library(fs)

# Confirm project root is resolved correctly
here::here()

# List all R scripts to verify naming convention
r_scripts <- dir_ls(here("R"), regexp = "\\.R$")
basename(r_scripts)
```

---

# Tool 2 — Version Control (Git)

## Rationale

Git records every change to the codebase with a cryptographic commit hash,
allowing exact reproduction of any historical state. Hosting the repository
on GitHub enables sharing, issue tracking, and DOI assignment for
permanent archiving. The commit hash at analysis time can be reported in a
methods section so readers can check out the exact code used.

## Git workflow applied in this project

The repository was initialised with `git init` before any analysis code was
written. All scripts were committed in a single initial commit; subsequent
debugging and race-harmonisation bug fixes were committed separately so each
commit corresponds to a logical, auditable change.

The `.gitignore` file below excludes raw data files (which may be large and
cannot be redistributed under cBioPortal terms) and generated output, keeping
the repository focused on reproducible *code* rather than derived artefacts.

```{r gitignore, eval=FALSE}
# Shown for documentation — actual file lives at project root

# Raw and processed data (may be large or sensitive)
data/raw/
data/processed/
data/*.rds
data/*.csv

# Output figures and tables (reproducible from code)
output/figures/
output/tables/

# R session artefacts
.Rhistory
.RData
.Rproj.user/
```

## Recording the commit hash at analysis time

The chunk below captures the commit hash and prints it so that any output
derived from this report is traceable to an exact code state.

```{r git-hash}
git_hash <- tryCatch(
  system("git rev-parse --short HEAD", intern = TRUE),
  error = function(e) "git not available"
)
cat("Analysis performed at commit:", git_hash, "\n")
cat("Analysis date:", format(Sys.Date(), "%Y-%m-%d"), "\n")
```

## Git commands used in this project

The four core commands from the slides cover all everyday operations:

```bash
# Initialise repository
git init

# Stage all R scripts and project files
git add R/ README.md .gitignore reproducibility_report.Rmd

# Commit with a descriptive message
git commit -m "Add complete MSK-CHORD 2024 racial disparities analysis pipeline"

# Push to GitHub (first time: set upstream)
git push -u origin master
```

---

# Tool 3 — Data Manipulation Practices

## Rationale

Dempsey & Morrison state: "Don't modify data in place. Avoid opening data in
Excel." This project follows three concrete rules derived from that principle:

1. **Raw data is never written to.** The two source files downloaded from
   cBioPortal (`msk_chord_2024_clinical_data.tsv` and `data_mutations.txt`)
   are read-only inputs; all transformations produce new objects or saved
   `.rds` files with distinct names.
2. **All data cleaning is scripted.** No manual editing in Excel was
   performed. Every filter, recode, and merge step is in `01_data_processing.R`
   and is fully auditable.
3. **The critical race-harmonisation step is explicitly documented** with
   a comment explaining the bug that was fixed — this is the kind of
   reasoning that disappears if changes are made interactively and not
   committed.

## Race/ethnicity harmonisation

This was the most consequential data-cleaning decision in the project.
The initial implementation used `grepl("Hispanic")` to assign the Hispanic/Latino
category. This misclassified approximately 5,000 White patients whose clinical
records contained the string "Non-Spanish; Non-Hispanic" (which contains
the substring "Hispanic"). The fix checks for "Non-Hispanic" before "Hispanic"
in the `case_when` priority order.

```{r race-harmonisation, eval=FALSE}
# CORRECT ordering in 01_data_processing.R
# ── The order of case_when clauses matters: ────────────────────────────────
# "Non-Spanish; Non-Hispanic" contains "Hispanic" as a substring.
# If we check grepl("Hispanic") first, ~5000 White patients are
# misclassified as Hispanic/Latino. Fix: check "Non-Hispanic" first.

df <- df %>%
  mutate(
    RACE_ETH = case_when(
      grepl("Non-Hispanic|non-hispanic", ETHNICITY, ignore.case = TRUE) &
        grepl("White|Caucasian",          RACE,      ignore.case = TRUE) ~ "White (non-Hispanic)",
      # ↑ Non-Hispanic White checked BEFORE generic Hispanic check
      grepl("Hispanic|Latino|Spanish",    ETHNICITY, ignore.case = TRUE) ~ "Hispanic/Latino",
      grepl("Asian",                      RACE,      ignore.case = TRUE) ~ "Asian",
      grepl("Black|African",              RACE,      ignore.case = TRUE) ~ "Black/African American",
      TRUE ~ "Other/Unknown"
    ),
    RACE_ETH = factor(RACE_ETH, levels = RACE_LEVELS)
  )
```

## Complete-case analysis and MICE imputation

Two analytical datasets are constructed from the cleaned data:

- **`nsclc_complete_case.rds`** — rows with no missing values in
  any covariate used in regression (n = 7,775). This is the primary dataset.
- **`nsclc_mice_imputed.rds`** — 20 multiply-imputed datasets created with
  `mice::mice()` for sensitivity analysis. MICE is applied *after* the
  complete-case file is saved, so a crash during imputation does not prevent
  downstream analyses from running.

```{r mice-strategy, eval=FALSE}
# From 01_data_processing.R
# ── Save complete-case BEFORE attempting MICE ──────────────────────────────
# This guarantees nsclc_complete_case.rds is available for all downstream
# scripts even if MICE fails (e.g., insufficient RAM or matrix rank issues).

saveRDS(nsclc_cc, file.path(DATA_DIR, "nsclc_complete_case.rds"))

# ── MICE: multiple imputation by chained equations ─────────────────────────
# m = 20 imputed datasets; maxit = 10 chained iterations per dataset.
# set.seed(42) in 00_setup.R ensures chains are fully reproducible.
mice_imp <- tryCatch(
  mice::mice(
    nsclc_mice_df, m = 20, maxit = 10,
    method = c(
      SMOKING_HISTORY = "logreg",  # binary outcome → logistic imputation
      STAGE           = "logreg",
      AGE             = "pmm"      # continuous → predictive mean matching
    ),
    printFlag = FALSE
  ),
  error = function(e) {
    message("MICE failed: ", e$message)
    NULL  # downstream scripts handle NULL gracefully
  }
)

if (!is.null(mice_imp))
  saveRDS(mice_imp, file.path(DATA_DIR, "nsclc_mice_imputed.rds"))
```

---

# Tool 4 — Programming Habits

## Rationale

The slides advise writing functions for any operation performed more than
once, and checking input parameters for type and known properties. This
project uses three design patterns that satisfy these requirements.

## Pattern 1: A single `source`-able setup file

`00_setup.R` is sourced at the top of every analysis script. It installs
missing packages, sets the global seed, defines directory paths, the
`DRIVER_GENES` character vector, and the `RACE_COLORS` named palette. This
eliminates copy-pasted boilerplate and guarantees every script uses identical
constants.

```{r setup-structure, eval=FALSE}
# 00_setup.R excerpt — sourced by every downstream script

set.seed(42)           # one seed, applied before any stochastic operation

DRIVER_GENES <- c("EGFR", "KRAS", "TP53", "STK11",
                  "KEAP1", "RB1", "ERBB2", "MET")

RACE_COLORS <- c(
  "White (non-Hispanic)"   = "#4477AA",   # colour-blind-safe palette
  "Asian"                  = "#EE6677",
  "Black/African American" = "#228833",
  "Hispanic/Latino"        = "#CCBB44",
  "Other/Unknown"          = "#BBBBBB"
)

DATA_DIR   <- here::here("data", "raw")   # no hard-coded paths
OUTPUT_FIG <- here::here("output", "figures")
OUTPUT_TAB <- here::here("output", "tables")
```

Note that `here::here()` constructs paths relative to the project root
regardless of the working directory at call time, satisfying the slides'
requirement to avoid `setwd()`.

## Pattern 2: Parameterised test and regression functions

Instead of repeating a `glm()` call eight times (once per gene), a single
function encapsulates the entire fitting and extraction logic. The function
checks for insufficient events (EPV < 10) before fitting, and falls back to
Firth penalised regression when standard GLM fails due to separation.

```{r logistic-function, eval=FALSE}
# From 04_logistic_regression.R

fit_logistic <- function(df, gene) {
  outcome <- paste0("MUT_", gene)
  if (!outcome %in% names(df)) return(NULL)   # input guard

  events <- sum(df[[outcome]], na.rm = TRUE)
  n_params <- length(strsplit(base_covars, "\\+")[[1]]) + 4
  epv <- events / n_params
  message(sprintf("  Gene: %-8s | events: %d | EPV ≈ %.1f", gene, events, epv))

  formula_str <- paste0(outcome, " ~ ", base_covars)

  fit <- tryCatch(
    glm(as.formula(formula_str), data = df,
        family = binomial(link = "logit"), na.action = na.omit),
    error = function(e) {
      message("Standard GLM failed for ", gene, ": ", e$message,
              " — trying Firth penalised regression")
      if ("logistf" %in% installed.packages()[, "Package"])
        logistf::logistf(as.formula(formula_str), data = df)
      else NULL
    }
  )
  fit
}

# Apply to all genes with a single purrr call
models_cc <- setNames(
  purrr::map(DRIVER_GENES, ~ fit_logistic(nsclc_df, .x)),
  DRIVER_GENES
)
```

The same modular pattern appears in `06_replication_cohorts.R` as
`run_logistic_cohort()`, which accepts a data frame and a cohort label and
returns a tidy results table — reused for both CRC and Breast cohorts without
code duplication.

## Pattern 3: Automatic test switching (chi-square vs Fisher)

The chi-square / Fisher switching logic in `03_chi_square_fdr.R` checks the
expected cell counts before every test and selects the appropriate method.
This prevents silent application of an invalid large-sample approximation
when rare mutations produce expected counts below five.

```{r chisq-switch, eval=FALSE}
# From 03_chi_square_fdr.R

test_gene_race <- function(df, gene) {
  tbl <- table(df$RACE_ETH, df[[paste0("MUT_", gene)]])

  expected   <- chisq.test(tbl)$expected
  use_fisher <- any(expected < 5)           # automatic test selection

  if (use_fisher) {
    res <- fisher.test(tbl, simulate.p.value = TRUE, B = 10000)
    test_used <- "Fisher (simulated)"
    stat <- NA_real_
  } else {
    res <- chisq.test(tbl)
    test_used <- "Chi-square"
    stat <- res$statistic
  }
  # ... extract p-value, Cramér's V, per-race prevalence
}
```

---

# Tool 5 — Pipeline Design

## Rationale

A pipeline formalises the sequence from raw data to final output. Without
one, it is easy to run scripts out of order, regenerate intermediate files
accidentally, or forget which parameter values were used. The slides describe
a pipeline as "a series of analysis steps that take you from data to output"
and note that even a README plus a directory of code is a valid starting point.

## Master pipeline: `07_run_all.R`

This project implements a sequential pipeline in a single R script. Each
step is represented as a list element with a path and a label. The `for`
loop sources each script in order, wrapping every `source()` call in
`tryCatch` so that a failure in one step (e.g., MICE memory overflow) prints
a clear error message and continues the pipeline rather than silently halting.

```{r pipeline-script, eval=FALSE}
# 07_run_all.R — complete listing

library(here)

steps <- list(
  list(script = "R/01_data_processing.R",    label = "Step 1: Data processing & imputation"),
  list(script = "R/02_descriptive.R",         label = "Step 2: Descriptive analysis & Table 1"),
  list(script = "R/03_chi_square_fdr.R",      label = "Step 3: Chi-square tests + BH-FDR"),
  list(script = "R/04_logistic_regression.R", label = "Step 4: Multivariable logistic regression"),
  list(script = "R/05_survival_analysis.R",   label = "Step 5: Survival analysis (KM + Cox PH)"),
  list(script = "R/06_replication_cohorts.R", label = "Step 6: Replication in CRC & Breast")
)

for (step in steps) {
  message("\n", paste(rep("=", 60), collapse = ""))
  message("▶ ", step$label)
  message(paste(rep("=", 60), collapse = ""))
  tryCatch(
    source(here::here(step$script)),
    error = function(e) {
      message("ERROR in ", step$script, ": ", e$message)
      message("Pipeline continuing with next step...")
    }
  )
}
```

## Pipeline data flow diagram

```
cBioPortal download
        │
        ▼
01_data_processing.R
   ├─ nsclc_complete_case.rds ──► 02, 03, 04, 05
   ├─ crc_complete_case.rds   ──► 06
   └─ breast_complete_case.rds──► 06
        │
        ▼
02_descriptive.R          → 02_table1_nsclc.csv
                            02_mutation_prevalence_*.pdf

03_chi_square_fdr.R       → 03_chisq_fdr_nsclc.csv
                            03_bubble_chisq_fdr.pdf

04_logistic_regression.R  → 04_OR_logistic_nsclc_cc.csv
                            04_forest_logistic_nsclc.pdf

05_survival_analysis.R    → 05_cox_hr_nsclc.csv
                            05_km_race_egfr_nsclc.pdf

06_replication_cohorts.R  → 06_OR_replication_cohorts.csv
                            06_replication_forest_all_cohorts.pdf
```

## Reproducing the analysis from scratch

A new investigator with access to cBioPortal can reproduce all results
with four commands:

```bash
# 1. Clone the repository
git clone https://github.com/weikai202/msk_chord_analysis.git
cd msk_chord_analysis

# 2. Download MSK-CHORD 2024 from cBioPortal:
#    https://www.cbioportal.org/study/summary?id=msk_chord_2024
#    Place the two files in data/raw/:
#      msk_chord_2024_clinical_data (1).tsv
#      data_mutations.txt

# 3. Run the entire pipeline (installs all packages automatically)
Rscript R/07_run_all.R

# 4. Knit this report
Rscript -e "rmarkdown::render('reproducibility_report.Rmd')"
```

---

# Key Results

This section loads the saved CSV outputs and displays the primary findings
to confirm that the pipeline ran end-to-end correctly.

## Step 3 output — Chi-square + BH-FDR results

The chunk below reads the saved association-test table. Six of eight driver
genes show statistically significant association with race/ethnicity after
Benjamini-Hochberg FDR correction (q < 0.05).

```{r chisq-results}
library(tidyverse)
library(here)
library(knitr)

chisq_df <- read_csv(here("output", "tables", "03_chisq_fdr_nsclc.csv"),
                     show_col_types = FALSE)

chisq_df %>%
  select(Gene, test_used, statistic, p_value, q_value, sig_label, cramers_v) %>%
  mutate(
    across(c(statistic, p_value, q_value, cramers_v), ~ round(.x, 4)),
    p_value  = formatC(p_value,  format = "e", digits = 2),
    q_value  = formatC(q_value,  format = "e", digits = 2)
  ) %>%
  kable(
    col.names = c("Gene", "Test", "Statistic", "p-value",
                  "q-value (BH)", "Sig.", "Cramér's V"),
    caption   = "Table 1. Unadjusted association between race/ethnicity and driver
                 gene mutation status (NSCLC, n = 7,775). BH-FDR at q < 0.05."
  )
```

EGFR shows the largest effect (Cramér's V = 0.229), with Asian patients
carrying a 53.9% mutation rate compared with 20.2% in White patients. KRAS
shows the second-largest effect (V = 0.144), driven by substantially lower
prevalence in Asian patients (9.8% vs. 31.2% in White).

## Step 4 output — Multivariable adjusted odds ratios

The chunk below loads the complete-case logistic regression results and
displays statistically significant associations after BH-FDR adjustment.

```{r logistic-results}
or_df <- read_csv(here("output", "tables", "04_OR_logistic_nsclc_cc.csv"),
                  show_col_types = FALSE)

or_df %>%
  filter(p_adj < 0.05) %>%
  mutate(across(c(OR, CI_low, CI_high), ~ round(.x, 3)),
         p_adj = formatC(p_adj, format = "e", digits = 2)) %>%
  select(Gene, Race_Group, OR, CI_low, CI_high, p_adj, sig_label) %>%
  arrange(Gene, Race_Group) %>%
  kable(
    col.names = c("Gene", "Race/Ethnicity", "Adj. OR",
                  "95% CI low", "95% CI high", "p (BH-adj)", "Sig."),
    caption   = "Table 2. Multivariable-adjusted odds ratios with FDR < 0.05.
                 Reference: White (non-Hispanic). Adjusted for smoking history,
                 stage, histologic subtype, sex, and age."
  )
```

After adjustment, EGFR enrichment in Asian patients strengthens (OR 2.93,
95% CI 2.43–3.53), confirming that the association is not explained by
differences in smoking, stage, or histology. KRAS and STK11 deficits in
Asian patients likewise persist after adjustment.

## Step 5 output — Cox proportional hazards model

The primary Cox model fits a `RACE_ETH × MUT_EGFR` interaction term. The
chunk below displays the main effects and interaction terms of primary
interest, excluding the nuisance HISTOLOGY coefficients.

```{r cox-results}
cox_df <- read_csv(here("output", "tables", "05_cox_hr_nsclc.csv"),
                   show_col_types = FALSE)

# Keep only RACE_ETH main effects, MUT_EGFR, key clinical covariates,
# and interaction terms
terms_keep <- c("RACE_ETHAsian",
                "RACE_ETHBlack/African American",
                "RACE_ETHHispanic/Latino",
                "RACE_ETHOther/Unknown",
                "MUT_EGFR",
                "SMOKING_HISTORYEver",
                "STAGEIV",
                "SEXMALE",
                "AGE",
                "RACE_ETHAsian:MUT_EGFR",
                "RACE_ETHBlack/African American:MUT_EGFR",
                "RACE_ETHHispanic/Latino:MUT_EGFR",
                "RACE_ETHOther/Unknown:MUT_EGFR")

cox_df %>%
  filter(term %in% terms_keep) %>%
  mutate(across(c(HR, CI_low, CI_high), ~ round(.x, 3)),
         p.value = formatC(p.value, format = "e", digits = 2)) %>%
  select(term, HR, CI_low, CI_high, p.value, sig_label) %>%
  kable(
    col.names = c("Term", "HR", "95% CI low", "95% CI high", "p-value", "Sig."),
    caption   = "Table 3. Multivariable Cox PH model with RACE_ETH × EGFR
                 interaction. Reference: White (non-Hispanic), EGFR wild-type.
                 Adjusted for smoking, stage, sex, age, histology."
  )
```

Stage IV carries the strongest prognostic signal (HR 2.88, p < 2×10⁻¹⁹⁶).
EGFR mutation is associated with modestly improved survival (HR 0.82,
p < 0.001). The interaction terms between race and EGFR status do not reach
statistical significance after adjustment, suggesting that the overall
racial survival disparity is not primarily mediated through differential
EGFR biology in this cohort.

---

# Reproducibility Checklist

The table below evaluates this project against the five tools from
Dempsey & Morrison (2023).

```{r checklist}
checklist <- tibble::tribble(
  ~Tool,                       ~Criterion,                                          ~Status,
  "1. File organisation",      "Descriptive, numbered script names",                "✓",
  "1. File organisation",      "Portable paths via here::here()",                   "✓",
  "1. File organisation",      "Raw data separated from outputs",                   "✓",
  "2. Version control",        "Git repository initialised",                        "✓",
  "2. Version control",        "Code hosted on GitHub",                             "✓",
  "2. Version control",        "Commit hash recorded at analysis time",             "✓",
  "2. Version control",        ".gitignore excludes data and outputs",              "✓",
  "3. Data manipulation",      "Raw data never modified in place",                  "✓",
  "3. Data manipulation",      "No Excel edits — all cleaning in R scripts",        "✓",
  "3. Data manipulation",      "Critical harmonisation logic documented",           "✓",
  "4. Programming habits",     "Functions for repeated operations",                 "✓",
  "4. Programming habits",     "Input guards in all modelling functions",           "✓",
  "4. Programming habits",     "No setwd() or hard-coded paths",                    "✓",
  "4. Programming habits",     "Global seed set in 00_setup.R",                     "✓",
  "5. Pipeline",               "Master script (07_run_all.R) runs all steps",       "✓",
  "5. Pipeline",               "tryCatch prevents silent failures",                 "✓",
  "5. Pipeline",               "Four-command reproduction from fresh clone",        "✓",
  "Dynamic report",            "This .Rmd integrates code, output, and narrative",  "✓"
)

kable(checklist,
      caption = "Reproducibility checklist evaluated against Dempsey & Morrison (2023).")
```

---

# Session Information

Recording the R session information at knit time ensures that any future
investigator can identify the exact package versions used in the analysis.
Version mismatches are a common source of non-reproducibility in R workflows.

```{r session-info}
sessionInfo()
```

---

# References

1. Dempsey, W. & Morrison, J. (2023). *Reproducible Research: Tools and
   Practices*. BIOSTAT 625, University of Michigan.
   (`jean997.github.io/rr_tools`)

2. Wilson, E. B. (1927). Probable inference, the law of succession, and
   statistical inference. *Journal of the American Statistical Association*,
   22(158), 209–212.

3. Benjamini, Y. & Hochberg, Y. (1995). Controlling the false discovery rate:
   a practical and powerful approach to multiple testing.
   *Journal of the Royal Statistical Society B*, 57(1), 289–300.

4. van Buuren, S. & Groothuis-Oudshoorn, K. (2011). mice: Multivariate
   Imputation by Chained Equations in R. *Journal of Statistical Software*,
   45(3), 1–67.

5. Schoenfeld, D. (1982). Partial residuals for the proportional hazards
   regression model. *Biometrika*, 69(1), 239–241.

6. MSK-CHORD 2024. Memorial Sloan Kettering Cancer Center Harmonized
   Oncology Research Database. cBioPortal for Cancer Genomics.
   https://www.cbioportal.org/study/summary?id=msk_chord_2024
